\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-02 paper/main.tex
Updates in Results 5

\subsection{Loss of cilia motility in specific subpopulation slows down transport}

We previously characterized three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the effects of these lineages on CSF flow and transport, 
we investigate region-specific cilia paralysis with the computational model.
Three cilia-covered regions were considered paralyzed separately:
the anterior ventricle, and the dorsal and ventral parts
of the middle ventricle (\Crefpanels{fig:fig5}{a}).
To simulate cilia paralysis, the cilia tangential traction boundary condition
is for each case changed to a free-slip boundary condition
in the paralyzed region,
prescribing zero tangential forces to
represent absence of cilia motion.

The CSF velocities are most strongly
affected when removing cilia in the dorsal part of the middle ventricle,
as the maximum velocity magnitude drops from 27.9 \textmu m/s to 7.9 \textmu m/s.
In vivo, these dorsal cilia generate the strongest flow at
this developmental stage of the fish~(\Crefpanels{fig:fig1}{c})~\cite{Olstad2019CiliaryDevelopment}.
In contrast, removing cilia in the anterior ventricle or the ventral region of the middle
ventricle has minimal impact, and results in maximum speeds
of 27.9 \textmu m/s and 27.4 \textmu m/s, respectively.
Of the total tangential forces exerted by the cilia, ventral cilia in the middle ventricle
contribute with 34\%, while the dorsal cilia exert 52\% of the total tangential
force~(\Crefpanels{fig:fig5}{b}). The magnitude of these force contributions are thus
more similar than the difference in their impact on the maximum velocity magnitudes.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{\textbf{a.} The three separate cilia regions for which cilia paralysis is simulated
    by imposing a zero tangential traction boundary condition in the respective region.
    \textbf{b.} The total tangential force applied to the boundary with the tangential traction
    boundary condition in each of the three separate ciliated regions, as well as the total force
    (the sum of the three regions), calculated with \Cref{eq:tangential_force_calculation}. 
    \textbf{c.} Mean concentration $\cbar(t)$ for the six regions of interest (ROIs) in the three 
    paralysis scenarios considered. The region named in the legend is the region
    where we modify the cilia tangential traction boundary condition to a free-slip condition
    in the CSF flow model.}
    \label{fig:fig5}
\end{figure}

Changes in the transport of Dendra2 are most prominent when removing the dorsal cilia
in the middle ventricle. All mean concentration profiles 
(except ROI 1, where we impose the photoconversion curve)
exhibit changes (\Crefpanels{fig:fig5}{c}).
Compared to the baseline model results, we observe less solute distribution
in the middle and anterior ventricles (ROIs 2--5),
and more rapid transport to the posterior ventricle (ROI 6).
On the other hand, removal of anterior cilia does not alter the 
mean concentrations (\Crefpanels{fig:fig5}{c}).
Finally, without ventral cilia in the middle ventricle,
more solute spreads towards the anterior ventricle (ROI 5) and
less solute spreads towards the posterior ventricle (ROI 6).

>>>>

\subsection{Regional loss of cilia motility delays transport}

We previously characterized three distinct ciliated cell lineages in
the zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment,
  DGama2025MotileBrain}.  To identify the effects of these lineages on
CSF flow and transport, we investigate region-specific cilia paralysis
with the computational model.  Three cilia-covered regions were
paralyzed separately: the anterior ventricle, and the dorsal and
ventral parts of the middle ventricle (\Crefpanels{fig:fig5}{a}).  To
simulate regional cilia paralysis, we remove the tangential traction exerted by
the cilia in the respective region.
%
%% the cilia tangential traction boundary condition is for each case
%% changed to a free-slip boundary condition in the paralyzed region,
%% prescribing zero tangential forces to represent absence of cilia
%% motion.
%

Removing cilia in the dorsal part of the middle ventricle has a strong
effect on the CSF velocities: the maximum velocity magnitude drops
from 27.9 \textmu m/s to 7.9 \textmu m/s. In vivo, these dorsal cilia
generate the strongest flow at this developmental stage of the
fish~(\Crefpanels{fig:fig1}{c})~\cite{Olstad2019CiliaryDevelopment}.
In contrast, only removing cilia in the anterior ventricle or the
ventral region of the middle ventricle has minimal impact, and results
in maximum speeds of 27.9 \textmu m/s and 27.4 \textmu m/s,
respectively.  We remark that out of the total tangential forces,
ventral cilia in the middle ventricle contribute by 34\%, while the
dorsal cilia exert 52\% of the total tangential
force~(\Crefpanels{fig:fig5}{b}). \mer{Thus, the CSF velocities are
  affected not only by the magnitude of the applied forces, but result
  from a nontrivial interplay between force, directionality and
  location.}
%% The magnitude of these force
%% contributions are thus more similar than the difference in their
%% impact on the maximum velocity magnitudes.

Turning to transport of Dendra2, we again note that removal of the
dorsal cilia has the most pronounced effect. All mean concentration
profiles (except ROI 1, where we impose the photoconversion curve)
change noticeably (\Crefpanels{fig:fig5}{c}). Compared to the baseline
model results, we observe less solute distribution in the middle and
anterior ventricles (ROIs 2--5), and more rapid transport to the
posterior ventricle (ROI 6). On the other hand, removal of anterior
cilia does not alter the mean concentrations
(\Crefpanels{fig:fig5}{c}). Finally, without ventral cilia in the
middle ventricle, more solute spreads towards the anterior ventricle
(ROI 5) and less solute spreads towards the posterior ventricle (ROI
6).
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{\textbf{a.} The three separate cilia regions for which cilia paralysis is simulated
    by imposing a zero tangential traction boundary condition in the respective region.
    \textbf{b.} The total tangential force applied to the boundary with the tangential traction
    boundary condition in each of the three separate ciliated regions, as well as the total force
    (the sum of the three regions), calculated with \Cref{eq:tangential_force_calculation}. 
    \textbf{c.} Mean concentration $\cbar(t)$ for the six regions of interest (ROIs) in the three 
    paralysis scenarios considered. The region named in the legend is the region
    where we modify the cilia tangential traction boundary condition to a free-slip condition
    in the CSF flow model.}
    \label{fig:fig5}
\end{figure}
